---
title: "LTER Working Group File Path Best Practices"
author: "NCEAS Scientific Computing Support Team"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NCEAS SciComp Team Best Practice Tips: **File Paths**

Welcome! We at the Scientific Computing Support Team have assembled this document to discuss our recommendations for handling **file paths**. When sharing code collaboratively (e.g., with GitHub) managing the difference between your folder structure and those of your colleagues can be tackled in a variety of ways. Failing to account for this can result in annoying errors where content are either not read in successfully or exported to the wrong folder. Below are our recommendations for file path management in a team context, we hope they help!

## 1) Preserve File Paths as Objects

First and foremost, we recommend that you begin each script by saving the paths to your files as objects. This makes it easy for each new user to remember that they need to alter those objects so that data are read in and saved out to the appropriate places.

```{r path_objects}
(my_path <- "path/to/my/file/")
```

## 2) Use `file.path()` to Make Path Objects

`file.path()` is a useful base R function that automatically accounts for the fact that Mac and PC operating systems use different slashes between folder names to indicate a path (one uses '/' and the other uses '\\')

```{r file.path_demo}
(my_path <- file.path("path", "to", "my", "file"))
```

Note that while you could use `paste(..., sep = '/')` instead, it does not account for the different slash between Mac and PC so `file.path() is preferable.

## 3) Use Path Objects when Reading/Writing Data

Now that you've saved your paths as objects using `file.path()` to account for operating system differences, you should use them during import/export steps! To do this, just wrap the argument (i.e., part of the function) that deals with the file name/path in another call to `file.path()` with the object you created earlier and the name of the file to import/export. See below for two examples:

```{r import, eval = F}
# Import
my_raw_data <- read.csv(file = file.path(my_path, "raw_data.csv"))

# Export
write.csv(x = data_obj, file = file.path(my_path, "tidy_data.csv"))
```

## Summary

We strongly recommend preserving your file paths as objects at the start of your scripts (using `file.path()`) to ensure that your scripts can be shared as easily as possible among your team without issues due to user-specific folder names or computer operating system interrupting the work that you set out to do.

Please feel free to reach out to NCEAS Programming Support (see our webpage [here](https://www.nceas.ucsb.edu/programming-support)) if you have any questions about this best practice manual and/or need help implementing some of this content.
